---
title: "454 project progress report 2"
author: "Tina Chen, Kyle Suelflow, Samina Stack"
date: "`r format(Sys.Date(),'%e %B, %Y')`"
output:
   prettydoc::html_pretty:
    theme: cayman
    toc: yes
    bibliography: Library.bib
    
abstract: XXX
---

```{r setup, include=FALSE, warning=FALSE}

knitr::opts_chunk$set(
  collapse = TRUE, 
  warning = FALSE,
  message = FALSE,
  fig.height = 2.75, 
  fig.width = 4.25,
  fig.align = 'center')

source('Cleaning.R')
```

## Introduction (Samina)
  * What is the motivation / goal behind this methodology? When is it useful?
  * clarify research goal: learn the methodology, then use data to illustrate it (talk about data later in data section)
  * In research goals, address why Bayesian spatial models might be preferable over non-Bayesian (i.e. motivate your exploration).
 
  

  
  

## Methodology (Tina)

#### Backrgound: Areal Data and neighborhood structure
Areal data differs from point data, which consists of measurements from a known set of geo-spatial points. The boundary of areas can be considered polygons determined by a closed sequence of ordered coordinates connected by straight line segments. 

The essential idea is that the probability of values estimated at any given location are conditional on the level of neighboring values. We assume that nearby locations tend to have similar characteristics or behaviors. Therefore, we need to define what is nearby locations.

For a set of $N$ areal units, the relationship between areal units is described by an $n \times n$ adjacency matrix $W$. The entries indicate whether two regions $n_i$ and $n_j$ are neighbors, with a value of 1 signifying adjacency and 0 indicating non-adjacency. It's worth noting that in models like Conditional Autoregressive (CAR) models, the neighbor relationship is symmetric, but a region is not considered its own neighbor ($W_{ii} = 0$).

  * What structure we are using here , for what reasons, present the demo 
  using Queen: If two polygons touch at all, even at one point, such as a corner, they are neighbors.
  
```{r}
# neighborhood structure
centroids <- st_centroid(st_geometry(simple_data_A), of_largest_polygon=TRUE)
W.nb <- poly2nb(simple_data_A, simple_data_A$grid_id) # using Queen
nb_Q_net <- nb2lines(nb = W.nb, coords = centroids, as_sf = TRUE)

simple_data_A %>%
ggplot() + 
  geom_sf(fill = 'white',color = 'darkgrey') + 
  geom_sf(data = centroids) + 
  geom_sf(data = nb_Q_net) + 
  theme_classic()

```
  
### Conditional Autoregressive (CAR) priors for spatial random effects


$\phi$, an n-length vector $\phi = (\phi_1, ..., \phi_n)^T$, is the spatial random variables that can represent spatial interactions between $n_i$ and $n_j$.

In the full conditional distribution, each $\phi_i$ is conditional on the sum of the weighted values of its neighbors, and has unknown variance.
$$
\begin{split}
 
 \phi_i | \phi_j ,j \neq i, 
 &\sim N (\sum\limits_{j=1}^n b_{ij} \phi_j, \tau_i^{-1})\\ 
 
\end{split}
$$
Where $\tau_i$ is a spatially varying precision parameter, and $b_{ii} = 0$.

The summation term in this expression is simply the weighted sum of the mean adjusted $\rho$ at all other locations $j$ - this may or may not be a reasonable assumption for a particular problem under consideration.

By Brook's Lemme, the joint distribution fo $\phi$ is then:
$$
\begin{split}
 \phi &\sim N(0, [D_{\tau}(I - \alpha B)]^{-1})\\
 \end{split}
$$

then the CAR prior specification simplifies to:

$$
\begin{split}
 \phi &\sim N(0, [\tau (D- \alpha W)]^{-1})\\
 \end{split}
$$


##### Assumption:

Where $W$ is the $n \times n$ adjacency matrix where entries $\{i,i\}$ are zero and the off-diagonal elements are 1 if regions $i$ and $j$ are neighbors and 0.

$D$ is the $n \times n$ diagonal matrix where entries $\{i,i\}$ are the number of neighbors of region $i$ and the off-diagonal entries are 0.

$D_{\tau} = \tau D$

$\alpha$ controls the amount of spatial correlation parameter that determines the size and nature (positive or negative) of the spatial neighborhood effect; $\alpha = 0$ implies spatial independence and $\alpha = 1$ implies complete spatial correlation. 

$B$ is the scaled adjacency matrix $D^{-1}W$

$I$ is an $n \times n$ identity matrix
 
When $\alpha$ is in the interval $(0,1)$, the covariance matrix $[D_{\tau}(I - \alpha B)]$ is positive definite, thus the joint distribution $\phi$ is proper.


###  Likelihood functionL: A Gaussian specification 

Suppose we have aggregated continuous data $y_1,y_2,...,y_n$ at $n$ locations, and we expect that neighboring locations will have similar data With a Gaussian likelihood:

$$
\begin{split}
y_i \sim \text{N}\left(X_i \beta + \phi_i \right)
\end{split}
$$

Where $X_i$ is a design vector (the $i^{th}$ row from a design matrix), $\beta$ is a vector of coefficients, $\phi_i$ is a spatial adjustment.
  
  


### Complete Bayesian Specification

Our posterior distribution is:
$$
\begin{split}
p(\phi, \beta, \alpha, \tau \mid y) \propto p(y \mid \beta, \phi) p(\phi \mid \alpha, \tau) p(\alpha) p(\tau) p(\beta)
\end{split}
$$

## Data analysis

research goal

  * An introduction the data context (who, what, where, why, when, how) and the data
  
  * An introduction to your research question and why it is important


#### data overview

  * smaller data points to demo
  
```{r}

pre_log <- simple_data_A %>% filter(POPGPW_2005_40 !=0) %>% 
ggplot(aes(x=MER2005_40))+
  geom_density()

post_log <-simple_data_A %>% filter(POPGPW_2005_40 !=0) %>% 
ggplot(aes(x=log(MER2005_40)))+
  geom_density()

ggarrange(pre_log, post_log, 
          nrow = 1, ncol = 2, labels = c("Pre log","Post log"))

pre_map <- mapview(simple_data_A, zcol = "MER2005_40")
real <- mapview(simple_data_A, zcol = "log(MER2005_40)")

leafsync::sync(pre_map,real, ncol = 2) 

```

```{r include=FALSE}
# ggpairs(data = simple_data_A, columns = c(3,9,12,13,14))
```


```{r}

# form <- `log(MER2005_40)`  ~ dis_to_water + log(POPGPW_2005_40) + TEMPAV_8008 + D3
# use simpler model
form <- `log(MER2005_40)`  ~ log(POPGPW_2005_40) 

lm_mod <- lm(formula=form, data = simple_data_A)
summary(lm_mod)

BIC(lm_mod)
summary(lm_mod)
simple_data_A$lm_pred <- predict(lm_mod)

simple_data_A %>% ggplot() +
  geom_sf(aes(fill = lm_pred)) +
  scale_fill_gradient2(mid = "white", high = "red", low = "blue") + 
  theme_classic()


# refined scale 

```

```{r}

# maran I test to test the spacial autocorrelation
W.list <- nb2listw(W.nb, style="B", zero.policy = TRUE) # list with spatial weights for W.nb

moran.mc(x=residuals(lm_mod), listw=W.list, nsim=1000)

```
The Moran’s I test has a p-value much less than 0.05, which suggests that the residuals contain substantial positive spatial autocorrelation.
  
  
```{r cache = true}
W <- nb2mat(W.nb, style="B")

# simulation
chain <- S.CARleroux(formula = form, data=simple_data_A, family="gaussian", W = W,
      burnin=500, n.sample=10000, n.chains=3, n.cores=3)


summary(chain)

# only has one beta 

plot(chain$samples$phi[,2:4][,1])
# plot(chain$samples$phi[,2:4][,3])
plot(chain$samples$beta[,2])

# plot(chain$samples$beta[,2:3][,1])
# plot(chain$samples$beta[,2:3][,2])
# plot(chain$samples$beta[,2:4][,3])
plot(chain$samples$rho)
plot(chain$samples$tau2)
# check all parameters

# saveRDS()

```

```{r}
y.fit <- chain$samples$fitted
y.fit <- t(t(y.fit))

simple_data_A$CAR <- apply(y.fit, 2, median)
simple_data_A$LL <- apply(y.fit, 2, quantile, 0.025)
simple_data_A$UL <- apply(y.fit, 2, quantile, 0.975)

simple_data_A$dis_to_water <- simple_data_A$dis_to_water
simple_data_A$pop <- simple_data_A$`log(POPGPW_2005_40)`
simple_data_A$temp <- simple_data_A$TEMPAV_8008
simple_data_A$GCP <- simple_data_A$`log(MER2005_40)`
simple_data_A$D3 <- simple_data_A$D3
simple_data_A$lm_pred <- simple_data_A$lm_pred

at <- seq(min(c(simple_data_A$CAR, simple_data_A$LL, simple_data_A$UL,simple_data_A$lm_pred)),
          max(c(simple_data_A$CAR, simple_data_A$LL, simple_data_A$UL,simple_data_A$lm_pred)),
          length.out = 8)


popuptable <- leafpop::popupTable(dplyr::mutate_if(simple_data_A,
                                  is.numeric, round, digits = 2),
zcol = c("pop","GCP", "temp","D3", "dis_to_water","LL", "UL","CAR"),
row.numbers = FALSE, feature.id = FALSE)


m1 <- mapview(simple_data_A, zcol = "CAR", map.types = "CartoDB.Positron",
              at = at, popup = popuptable)
m2 <- mapview(simple_data_A, zcol = "LL", map.types = "CartoDB.Positron",
              at = at, popup = popuptable)
m3 <- mapview(simple_data_A, zcol = "UL", map.types = "CartoDB.Positron",
              at = at, popup = popuptable)

lm_model <- mapview(simple_data_A, zcol = "lm_pred", map.types = "CartoDB.Positron",
              at = at, popup = popuptable)

real <- mapview(simple_data_A, zcol = "log(MER2005_40)", map.types = "CartoDB.Positron",
              at = at, popup = popuptable)


```





```{r}
# INLA


nb2INLA("map.adj", W.nb)
g <- inla.read.graph(filename = "map.adj")

formula <- `log(MER2005_40)` ~ log(POPGPW_2005_40) 
  f(re_u, model = "besag", graph = g, scale.model = TRUE)
  
res <- inla(formula, family = "gaussian", data = simple_data_A,
control.predictor = list(compute = TRUE),
control.compute = list(return.marginals.predictor = TRUE))

res$summary.fixed
summary(res$summary.fitted.values)

# Posterior mean and 95% CI
simple_data_A$PMINLA <- res$summary.fitted.values[, "mean"]
# simple_data_A$LLINLA <- res$summary.fitted.values[, "0.025quant"]
# simple_data_A$ULINLA <- res$summary.fitted.values[, "0.975quant"]

minla <- mapview(simple_data_A, zcol = "PMINLA", map.types = "CartoDB.Positron",
              at = at, popup = popuptable)
```

```{r}
# all map
m <- leafsync::sync(m1,lm_model, real,minla, ncol = 2) 

m
```
